---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r load}
library(mice)
library(lattice)
data <- read.csv('activity.csv')
```

## Imputing missing values

Due to practical concerns the imputing step is done first (you can not calculate mean otherwise).
Imputing is done by mice library, because it is a little more sophisticated than just taking the mean.

```{r}
# check number of NA's:
md.pattern(data)

impute <- mice(data, m=1, maxit = 10, method = 'pmm', seed = 500)
imputed_data <- complete(impute, 1)
```

## What is mean total number of steps taken per day?

```{r}
data$date <- as.Date(data$date, format='%Y-%m-%d')
imputed_data$date <- as.Date(imputed_data$date, format='%Y-%m-%d')
per_day_mean <- aggregate(data, list(period=data$date), mean)
per_day_mean_imp <- aggregate(imputed_data, list(period=imputed_data$date), mean)

per_day_median <- aggregate(data, list(period=data$date), median)
per_day_median_imp <- aggregate(imputed_data, list(period=imputed_data$date), median)
# Mean steps per day:
print('Mean steps per day (without impution):')
print(per_day_mean[, 1:2])
print('Mean steps per day (with impution):')
print(per_day_mean_imp[, 1:2])
print('Median steps per day (without impution):')
print(per_day_median[, 1:2])
print('Median steps per day (with impution):')
print(per_day_median_imp[, 1:2])

# total mean
sum(per_day_mean_imp$steps)
# total median
sum(per_day_median_imp$steps)

hist(per_day_mean[,2], main='Histogram of mean steps per day (no impution)', xlab='Mean steps per day (no impution)')

hist(per_day_mean_imp[,2], main='Histogram of mean steps per day (with impution)', xlab='Mean steps per day (with impution)')
```
## What is the average daily activity pattern?

Plotted using Lattice plotting system.

```{r}
# Add weekdays and weekend variable
imputed_data$day <- as.factor(ifelse((weekdays(imputed_data$date) == "zaterdag" | weekdays(imputed_data$date) == "zondag"), 'weekend', 'weekday'))

# Average over time series
avg_split <- aggregate(imputed_data$steps, by= list(imputed_data$interval, imputed_data$day), mean)
avg_total <- aggregate(imputed_data$steps, by= list(imputed_data$interval), mean)

plot(avg_total, type = "l", xlab = "Time Intervals (5-minute)", 
    ylab = "Mean number of steps taken (all Days)", main = "Average Steps Taken at 5 minute Intervals", 
    col = "blue")

names(avg_split) <- c("interval", "day", "steps")
xyplot(steps ~ interval | day, data = avg_split, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
```

## Are there differences in activity patterns between weekdays and weekends?

Weekdays seem to have a more concentrated area where steps are being taken. It is likely that this is between 'working hours'. Weekend days have a more random pattern.